---
title: "Practical Machine Learning Assignment"
author: "tc"
date: "Monday, April 20, 2015"
output: html_document
---
**question:**
Can I predict "which" activity was performed at a specific point in time?
More specifically, Can I predict the manner (specified execution or common 
mistake) in which the test subjects did the exercise. 

**Input data:**
Unilateral Dumbbell Biceps Curl in five different fashions: "classe" variable
exactly according to the specification (Class A),: specified execution 
throwing the elbows to the front (Class B), : common mistake
lifting the dumbbell only halfway (Class C), : common mistake
lowering the dumbbell only halfway (Class D) : common mistake
and throwing the hips to the front (Class E).: common mistake

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

Loading libraries:
```{r, echo=TRUE}
library(ggplot2)
library(lattice)
library(caret)
library(randomForest)
library(rpart)
library(rattle)
```

Loading data
```{r, echo=TRUE}
pml_testing <-read.csv("pml-testing.csv")
pml_training <-read.csv("pml-training.csv", na.strings= c("", "NA"), header = TRUE)
dim(pml_testing); dim(pml_training)
```

Set seed for reproducibility
```{r, echo=TRUE}
set.seed(32343)
```

Removing columns from training dataset that contain blanks or NA
```{r, echo=TRUE}
sub_pml_training <- pml_training[ , ! apply(pml_training , 2 , function(x) any(is.na(x)))]
dim(sub_pml_training) 
```

Removing the first seven columns from training dataset (number, username, time)
as the first seven columns are unnecessary for prediction
```{r, echo=TRUE}
sub1_pml_training <- sub_pml_training[,-(1:7),drop=FALSE] 
dim(sub1_pml_training)
```

Creating partition for testing
```{r, echo=TRUE}
inTrain <- createDataPartition(y=sub1_pml_training$classe,
                p=0.6, list=FALSE)
training <- sub1_pml_training[inTrain,]
testing <- sub1_pml_training[-inTrain,]
dim(training)
dim(testing)
```

Model:
```{r, echo=TRUE}
model_rf <- randomForest(classe ~ ., data = training)
model_rf
```
Random forest/caret does not require cross validation as it is done internally

Other Models:
1. Tried using the method rpart but predictions were not as expected for the testing 
set (3000+ in order for each classe letter):
```{r, eval=FALSE}
modFit <- train(classe ~., data=training, method="rpart")
print(modFit$finalModel)
fancyRpartPlot(modFit$finalModel)
predict(modFit, newdata=pml_testing)
```
2. Tried using the method rf but didn't have enough computer memory:
```{r, eval=FALSE}
modFitrf <- train(classe ~., data=training, methods="rf", prox=TRUE)
```
Would like to try more models - but running out of time to knit and submit.

Predicting testing set
```{r, echo=TRUE}
pml_test_prediction <- predict(model_rf, newdata=testing)
confusionMatrix(pml_test_prediction, testing$classe)
```
Accuracy is 99.5%  
Out of sample error is 1 - accuracy = 0.5%  
OOB estimate of  error rate: 0.54%  

Predicting 20 different test cases. (don't print answers)
```{r, echo=TRUE}
classepredict = predict(model_rf, newdata=pml_testing)
```

To submit the test results I used: 
```{r, eval=FALSE}
pml_write_files = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
}
answers = predict(model_rf, newdata=pml_testing)
pml_write_files(answers)
```
**All predictions were correct.** 

Thanks as always to the contributors of stackedoverflow for many helpful codes.   
Thanks to the forum for many helpful hints and answers.
